---
title: "Chocolate Ratings Analysis"
description: |
  Analysis of Chocolate Ratings data from TidyTuesday
author: Arsene Laurent
date: February 19, 2025
format: html
---

```{r}
#| echo: false
#| message: false
tuesdata <- tidytuesdayR::tt_load('2022-01-18')
tuesdata <- tidytuesdayR::tt_load(2022, week = 3)

chocolate <- tuesdata$chocolate
```

```{r}
#| echo: false
#| message: false
library(ggplot2)
library(dplyr)
library(ggthemes)
library(forcats)

avg_ratings <- chocolate |>
  group_by(company_location) |>
  summarize(average_rating = mean(rating)) |>
  arrange(desc(average_rating))

top_15 <- avg_ratings |>
  slice(1:15)

top_15$company_location <- factor(top_15$company_location, levels = top_15$company_location)

ggplot(top_15, (aes(x = company_location, y = average_rating))) +
  geom_bar(stat = "identity") +
  labs(x = "Country", y = "Avg Rating", title = "Top 15 Countries by Average Rating") +
  theme_solarized(light = FALSE) +
  theme(plot.margin = margin(1, 1, 3, 1, "cm"), axis.text.x = element_text(angle = 45, hjust = 1))
```


This data comes from the 2022-01-18 folder of the TidyTuesday GitHub repository. After selecting the relevant variables and creating a graph with rating as our y-axis and Country on the x-axis, we can see that companies based in Chile had the highest average rating for their chocolate bars. Unfortunately, this data is highly skewed by the fact that there were only two Chilean companies who had their chocolate bars rated. Below is a graph plotting all of the data, showing which country's companies had the highest rated chocolate bars.

```{r}
#| echo: false
#| message: false
library(ggplot2)
library(dplyr)
library(ggthemes)
library(forcats)

chocolate <- chocolate |> 
  mutate(cocoa_percent = as.numeric(gsub("%", "", cocoa_percent)))

by_percent <- chocolate |>
  arrange(desc(cocoa_percent))
ggplot(by_percent, (aes(x = cocoa_percent, y = rating))) +
  geom_point(aes(color = company_location)) +
  labs(x = "Cocoa Percent", y = "Rating", title = "Rating by Cocoa Percent") +
  theme_solarized(light = FALSE) +
  theme(
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.5, "cm"),
    plot.margin = margin(1, 1, 2, 1, "cm"),
    legend.position = "right", plot.title = element_text(vjust = 4))
```
Unfortunately, this plot doesn't offer us much more information on which country produces the "best" chocolate, though we can observe that the highest rated chocolate is pretty consistently within the 60-80$%$ range of cocoa concentration.


Here is a link to the original data by Brady Brelinski: https://flavorsofcacao.com/chocolate_database.html  